Towards user-friendliness in proof assistants: an effect-based attempt

### 2020-09-11 paper/main.tex
more text and better infographic

\author{April Gonçalves}
\email{april@cyberglot.me}
\affiliation{
  \institution{Metastate AG}
  \city{Edinburgh}
  \state{UK}
}

@@

In this section, we introduce our tool, named Witch.

\paragraph{Downsides of our approach}
There are two main styles for writing proofs within a proof assistant: external
verification and internal verification. The latter concerns itself with code
that is \textit{correct-by-construction}, meaning that only valid states allowed and it
carries all information necessary to make proofs trivial; while the former
concerns itself with proofs about non-proof-carrying code.
Witch's primarily focus is on external verification. We believe that newcomers
will struggle to write proof-aware code, since it departs strongly from
mainstream programming, even within the functional style. We expect users to
write their code as they normally would do and then prove necessary properties.
However, Witch does not support writing correct-by-construction code.

@@

>>>>

\author{April Gonçalves}
\email{april@metastate.dev}
\affiliation{
  \institution{Metastate AG}
  \city{Edinburgh}
  \state{UK}
}

@@

In this section, we introduce our tool, named Witch\footnote{Our tool's name is
  a play after assistant tools colloquially called ``wizards''. There is no
  consensus of what a wizard is or what are exactly the tasks it is supposed
  to help users. It seems to be used mostly for multiple-step and/or
  configuration features, however. We went for the name ``witch'' to align it to
the idea of assistant tools, while dodging an overloaded terminology.}.

\paragraph{Downsides of our approach}
There are two main styles for writing proofs within a proof assistant: external
verification and internal verification. The latter concerns itself with code
that is \textit{correct-by-construction}, meaning that only valid states allowed and it
carries all information necessary to make proofs trivial; while the former
concerns itself with proofs about code that does not carry any additional
information for proving purposes.
Witch's primarily focus is on external verification. We believe that newcomers
will struggle to write proof-aware code, since it departs strongly from
mainstream programming, even within the functional style. We expect users to
write their code as they normally would do and then prove necessary properties.
However, Witch does not support writing correct-by-construction code. As proofs
are erased during compilation-time in Juvix, Witch could be extended
to be indexed by proof objects and parameterised over any type, such as
\texttt{\{x y: a\} -> Witch (x * y ≡ y * x) a} where we make sure that \texttt{*}
over \texttt{a} is commutative. However, it is unclear whether having all user's code
under a monad-like structure would lead to well-structured code.

@@

\section{Conclusion and Future Work}

While it is hard to say whether we have achieve our goals before user
evaluations, Witch steps towards a future where proof assistants are more
friendly towards users, particularly newcomers.
